Visualizing Multivariate PDFs

### 2016-09-22 sections/draft.tex
papers, rearranging, etc

Some research institutions provide a tool to explore data they created, such as the \cite{src:esrlpsd}, but since all these tools are site specific, a researcher would either have to use the datasets on the website or submit their datasets for uploading 
(assuming their dataset meets the submission criteria for an aggregation site). A scientist also has the option of using the CDAT suite of libraries, developed by the \cite{WilliamsEtAl13}, GrADS, \cite{src:grads},and Ferret, \cite{src:HankinEtAl96}. The major critique with all of these though is that they mostly render the data as is, with limited support for statistical analysis and the application of typical meterological analysis. Support for even common machine learning techniques often requires extending the library; therefore for the purpose of this discussion it is somewhat useful to treat these tools more as plotting libraries than visualization systems. %%somewhere scope out/differentiate between plotting and viz

>>>>

\subsection{Structure & Seperability}

\subsection{Outlier Detection}
Some research institutions provide a tool to explore data they created, such as the \cite{src:esrlpsd}, but since all these tools are site specific, a researcher would either have to use the datasets on the website or submit their datasets for uploading 
(assuming their dataset meets the submission criteria for an aggregation site). A scientist also has the option of using the CDAT suite of libraries, developed by the \cite{WilliamsEtAl13}, GrADS, \cite{src:grads},and Ferret, \cite{src:HankinEtAl96}. The major critique with all of these though is that they mostly render the data as is, with limited support for statistical analysis and the application of typical meterological analysis. Support for even common machine learning techniques often requires extending the library; therefore for the purpose of this discussion it is somewhat useful to treat these tools more as plotting libraries than visualization systems. %%somewhere scope out/differentiate between plotting and viz

### 2016-09-22 sections/section1.tex
papers, rearranging, etc

\begin{document}
\section{Seperable Dimensions}
\label{sec:sepdim}

While bar graphs and histograms start giving a sense of the variability in
data, they are limited to showing the dispersion of the data only in 1
dimension; namely the measurements themselves. In their survey of the history
of the boxplot \cite{wickham2011} Wickham and Stryjewski trace the evolution of
the boxplot from a difinitive box to amorphous envelopes.  They start by laying out
the core statistics of a boxplot: a box centered at the median (or mean) and
bound by the upper and lower quartiles (termed the hinges), and two whiskers which represent the
extremes (often 1.5 the outer quartiles, sometimes the $\alpha$ bounds). Sometimes there's
also a dot to signify outliers, as seen in \ref{fig:boxplot}, but one of
the strengths of the box plot has been in it's flexibility.

\subsection{Discrete observations}

@@

\begin{figure}
  \includegraphics[width=.75\linewidth]{contour_weather}
  \caption{}
\label{fig:countour}
\end{figure}
While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker2013} aim instead to capture the variablity of the
uncertainity via a band depth ordering metric. Each ensemble members band depth
is computed as sum of the probabilities that the observations in any given
ensemble fall within the max\-min envelope defined by any two other
ensembles. The bands are then sorted by band depth such that the median is the
ensemble with about 50\% of its members withen all envelopes formed by other
bands (so most centered). Outer bands are chosen according to the task at
hand. In \cite{whitaker2013}, they apply the countour boxplots to temperature
visualizations. The authors argue that countour boxplots are an improved visual
idiom over the traditional plots seen in Figure~\ref{fig:spag} because
as seen in , the spagehetti plots get nosy as the number of plots increase
and so it's hard to tease out specific patterns. Instead in Figure~\ref{fig:countour}, the authors
remove most of the bands and instead visualize an outlier envelope(light gray)
and a more central envelope(dark gray). It retains much of the information of
the spagethhit, but removes the visual noise of the lines.

>>>>

\begin{document}

\section{Independent Observations}
\label{sec:independent}

\subsection{Univariate Observations}

%%Histogram paper + chambers + kde/pdf/density estimaton

\subsection{Multivariate Observations}
While bar graphs and histograms start giving a sense of the variability in
data, they are limited to showing the dispersion of the data only in 1
dimension; namely the measurements themselves. In their survey of the history
of the boxplot \cite{wickham2011} Wickham and Stryjewski trace the evolution of
the boxplot from a difinitive box to amorphous envelopes.  They start by laying out
the core statistics of a boxplot, which can also be seen in Figure~\ref{fig:boxplot}.
\begin{description}
  \item[Median (Q2)] 50\% quartile
  \item[Inter Quartile Range (IQR)]
\end{description}

\begin{tabular}{|r| r |r|r|}
 & Lower & Upper\\
  \multicolumn{2}{Hinge} & Q1: 25\%  & Q3: 75\% \\
  \multicolumn{2}{Interquartile range} & \multicolumn{2}{IQR = Q3-Q1}\\
  \multirow{2}{Fence} & Inner & Q1-1.5*IQR & Q3-1.5*IQR \\
  & outer & Q1-3*IQR & Q3+3*IQR \\
\end{tabular}

%%Insert a labeled boxplot with fences! (this should probably just be an
%%annotated figure

\subsection{Multivariate Observations}

@@

\begin{figure}
  \includegraphics[width=.75\linewidth]{contour_weather}
  \caption{\label{fig:countour}}
\end{figure}

While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker2013} aim instead to capture the variablity of the
uncertainity via a band depth ordering metric. Each ensemble members band depth
is computed as sum of the probabilities that the observations in any given
ensemble fall within the max-min envelope defined by any two other
ensembles. The bands are then sorted by band depth such that the median is the
ensemble with about 50\% of its members withen all envelopes formed by other
bands (so most centered). Outer bands are chosen according to the task at
hand. In \cite{whitaker2013}, they apply the countour boxplots to temperature
visualizations. The authors argue that countour boxplots are an improved visual
idiom over the traditional spagehtti plots (%insert citation for them...)
because the spagehetti plots get nosy as the number of plots increase
and so it's hard to tease out specific patterns. Instead in Figure~\ref{fig:countour}, the authors
remove most of the bands and instead visualize an outlier envelope(light gray)
and a more central envelope(dark gray). It retains much of the information of
the spagethhit, but removes the visual noise of the lines.
